---
title: 'Part 1: binary logistic regression'
author: "Roshni Katrak-Adefowora"
date: "1/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(jtools)
```

```{r}
penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color=species))
```

Create subset with only Adelie and Chinstrap penguins
```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>%  #levels(adelie_chinstrap$species) in console tells you there are 3 factors -- Gentoo is still there (because it is a factor, even if there are no observations)
  mutate(species = fct_drop(species)) #drop factors that have no observations

#now levels(adelie_chinstrap$species) should work
```

Let's just dataviz of those two species
```{r}
ggplot(data=adelie_chinstrap, aes(x=body_mass_g, y=flipper_length_mm))+
  geom_point(aes(color=sex))+
  facet_wrap(~species)

#looks like theres some size difference, will binomial logistic regression capture that?
```

Binary logistic regression

variables:
- body mass
- flipper length
- penguin sex

```{r}
adelie_chinstrap_blr <- glm(species ~ body_mass_g + flipper_length_mm + sex, 
                            data=adelie_chinstrap,
                            family="binomial")
```

Take a look at the outcomes
```{r}
adelie_chinstrap_blr
summary(adelie_chinstrap_blr) #null hypothesis: value of coeff = 0

#levels(adelie_chinstrap$species) Adelie shows up first and Chinstrap second so Adelie is coded as zero and Chinstrap as 1. So in log linear expression, its gonna be log odds (left side) of Chinstrap

#as body mass inc, odds of penguin being a Chinstrap dec (although non-significant, but that's what the sign indicates)

#compared to female, if a penguin is male it is less likely to be Chinstrap penguin (doesn't make much sense for this specfic categorical variable)

#as flipper length inc, odds of penguin being Chinstrap inc (significant - does this make sense?)
```

```{r}
ggplot(data=adelie_chinstrap, aes(x=species, y=flipper_length_mm))+
  geom_jitter(aes(color=sex))
#positive coef does align with graph
```

Let's look at actual probabilities (since log odds are hard to understand)

What are the actual probabilities of being Chintrap for each of the existing observations in our adelie_chinstrap dataframe
```{r}
#add columns to check how well model fits the data!
#fitted column: based on blr, the probability (% chance) based on body mass, flipper length, and sex, that this penguin is a Chinstrap. Some will be wrong. We can then go on to check how often the model is correctly classifying each species.
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict="response")
```

Plotting fitted probabilities
```{r}
ggplot(data=blr_fitted, aes(x=flipper_length_mm, y=.fitted))+
  geom_point(aes(color=sex, pch=species))+ #pch for different types of points for each species
  geom_smooth(aes(color=sex), se=FALSE)+
  labs(x="Flipper length (mm)",
       y="Probability of outcome Chinstrap")

#shows inc probability of being Chinstrp with inc flipper length

#doesn't show actual outcomes, but that can be added
```

Plot model based on different variables
```{r}
#Quick overview of how probability changes wrt a model variable
effect_plot(adelie_chinstrap_blr,
            pred=flipper_length_mm, #predictor to be major axis
            y.label = "Probability of Chinstrap",
            interval = TRUE) #add CI

effect_plot(adelie_chinstrap_blr,
            pred=body_mass_g, #predictor to be major axis
            y.label = "Probability of Chinstrap",
            interval = TRUE) #add CI
```

What's the probability that a female penguin weighing 3410g with a flipper length of 192mm is a Chinstrap?
```{r}
ex_1 <- predict(adelie_chinstrap_blr,
                data.frame(sex="female", #dataframe with values under same variable names
                           body_mass_g = 3410,
                           flipper_length_mm = 192),
                type = "response") #give probability, not log odds (default) 

ex_1
```

